---
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
---

```{r intro, echo=FALSE, results="hide"}
knitr::opts_chunk$set(echo=FALSE, 
                      message=FALSE, 
                      comment = "", 
                      warning=FALSE, 
                      results="asis") 

date<-as.Date(Sys.time(	), format='%d%b%Y')

suppressWarnings(suppressMessages(library(dplyr)))
suppressWarnings(suppressMessages(library(plyr)))
suppressWarnings(suppressMessages(library(tidyr)))
suppressWarnings(suppressMessages(library(tidyverse)))
suppressWarnings(suppressMessages(library(janitor)))

suppressWarnings(suppressMessages(library(readxl)))
suppressWarnings(suppressMessages(library(plotly)))
suppressWarnings(suppressMessages(library(writexl)))
suppressWarnings(suppressMessages(library(stringr)))
suppressWarnings(suppressMessages(library(RColorBrewer))) 
```

# Import
```{r directory}
mydir<-c("~/Dropbox/0Kids/0iSquareed_CollegeSoup/")

opecss<-c(paste0(mydir, "OPE_CSS/Custom/"))
#dir(opecss)

filenames <- list.files(path = opecss, pattern = ".csv$")
```

```{r import_arrests}
locations<-c("On_campus", 
             #"On_campus_Student_Housing_Facilities", 
             "Noncampus",
             "Public_Property")

dta<-read_csv(paste0(opecss,"Arrests_", locations[1], ".csv")) 
    
    dim(dta)
    dta$location <- locations[1]
    dim(dta)
    
for(i in 2:length(locations)){
    dtatemp<-read_csv(paste0(opecss,"Arrests_", locations[i], ".csv")) 
    
    print(dim(dtatemp[i]))
    dtatemp$location <- locations[i]
    dta<-rbind(dta, dtatemp)
    dim(dta)
    #print(str(dta[i]))
}
    
colnames(dta)<-tolower(colnames(dta))
dta<-clean_names(dta)

dta_arrests<-dta
```

```{r import_criminal}
locations<-c("On_campus", 
             #"On_campus_Student_Housing_Facilities", 
             "Noncampus",
             "Public_Property")

dta<-read_csv(paste0(opecss,"Criminal_Offenses_", locations[1], ".csv")) 
    
    dim(dta)
    dta$location <- locations[1]
    dim(dta)
    
for(i in 2:length(locations)){
    dtatemp<-read_csv(paste0(opecss,"Criminal_Offenses_", locations[i], ".csv")) 
    
    print(dim(dtatemp[i]))
    dtatemp$location <- locations[i]
    dta<-rbind(dta, dtatemp)
    dim(dta)
    #print(str(dta[i]))
}
    
colnames(dta)<-tolower(colnames(dta))
dta<-clean_names(dta)

dta_criminal<-dta
```

```{r import_disciplinary}
# locations<-c("On_campus", 
#              "On_campus_Student_Housing_Facilities", 
#              "Noncampus",
#              "Public_Property")
locations<-c("On_campus",
             #"Student_Housing_Facilities",
             "Noncampus",
             "Public_Property")

dta<-read_csv(paste0(opecss,"Disciplinary_Actions_", locations[1], ".csv")) 
    
    dim(dta)
    dta$location <- locations[1]
    dim(dta)
    
for(i in 2:length(locations)){
    dtatemp<-read_csv(paste0(opecss,"Disciplinary_Actions_", locations[i], ".csv")) 
    
    print(dim(dtatemp[i]))
    dtatemp$location <- locations[i]
    dta<-rbind(dta, dtatemp)
    dim(dta)
    #print(str(dta[i]))
}
    
colnames(dta)<-tolower(colnames(dta))
dta<-clean_names(dta)

dta_disciplinary<-dta
```

```{r import_hate}
locations<-c("On_campus", 
             #"On_campus_Student_Housing_Facilities", 
             "Noncampus",
             "Public_Property")

dta<-read_csv(paste0(opecss,"Hate_Crimes_", locations[1], ".csv")) 
    
    dim(dta)
    dta$location <- locations[1]
    dim(dta)
    
for(i in 2:length(locations)){
    dtatemp<-read_csv(paste0(opecss,"Hate_Crimes_", locations[i], ".csv")) 
    
    print(dim(dtatemp[i]))
    dtatemp$location <- locations[i]
    dta<-rbind(dta, dtatemp)
    dim(dta)
    #print(str(dta[i]))
}
    
colnames(dta)<-tolower(colnames(dta))
dta<-clean_names(dta)

dta_hate<-dta
```

```{r import_vawa}
locations<-c("On_campus", 
             #"On_campus_Student_Housing_Facilities", 
             "Noncampus",
             "Public_Property")

dta<-read_csv(paste0(opecss,"VAWA_Offenses_", locations[1], ".csv")) 
    
    dim(dta)
    dta$location <- locations[1]
    dim(dta)
    
for(i in 2:length(locations)){
    dtatemp<-read_csv(paste0(opecss,"VAWA_Offenses_", locations[i], ".csv")) 
    
    print(dim(dtatemp[i]))
    dtatemp$location <- locations[i]
    dta<-rbind(dta, dtatemp)
    dim(dta)
    #print(str(dta[i]))
}
    
colnames(dta)<-tolower(colnames(dta))
dta<-clean_names(dta)

dta_vawa<-dta
```

```{r check_imported_data}
dim(dta_arrests)
dim(dta_criminal)
dim(dta_disciplinary)
dim(dta_hate)
dim(dta_vawa)

table(dta_arrests$location)
table(dta_criminal$location)
#table(dta_disciplinary$location)
#table(dta_hate$location)
table(dta_vawa$location)

colnames(dta_arrests)
colnames(dta_criminal)
#colnames(dta_disciplinary)
#colnames(dta_hate)
colnames(dta_vawa)

```        

# Gen var
```{r genvar_arrests, results='hide'}
temp<-dta_arrests

# temp<-temp%>%
#     filter(institution_name=="Johns Hopkins University")%>%
#     arrange(institution_name, opeid, campus_name, campus_id, survey_year )%>%
#     filter(campus_id<=2)%>%
#     filter(survey_year>=2017)
# 
# dim(temp)
dta <- temp%>%
    select(institution_name, opeid, campus_name, campus_id, survey_year, 
             institution_size, colnames(temp[9: ncol(temp)-1]))%>%
    group_by(institution_name, opeid, campus_name, campus_id, survey_year, 
             institution_size)%>%
    summarise_all(list(sum), na.rm = TRUE)
# dim(dta)

dta$total_arrests <- rowSums(dta[ , colnames(dta[ , 7:ncol(dta)])])

dta_arrests <- dta
```    

```{r genvar_criminal, results='hide'}
# temp<-dta_criminal%>%
#     filter(institution_name=="Johns Hopkins University" &
#                 campus_name=="JHU  HOMEWOOD  CAMPUS- BALTIMORE- MARYLAND")
temp<-dta_criminal

dta <- temp%>%
    select(institution_name, opeid, campus_name, campus_id, survey_year, 
             institution_size, colnames(temp[9: ncol(temp)-1]))%>%
    group_by(institution_name, opeid, campus_name, campus_id, survey_year, 
             institution_size)%>%
    summarise_all(list(sum), na.rm = TRUE)

dta$total_criminal <- rowSums(dta[ , colnames(dta[ , 7:ncol(dta)])])

dta_criminal <- dta

```    

```{r genvar_vawa, results='hide'}
temp<-dta_vawa

dta <- temp%>%
    select(institution_name, opeid, campus_name, campus_id, survey_year, 
             institution_size, colnames(temp[9: ncol(temp)-1]))%>%
    group_by(institution_name, opeid, campus_name, campus_id, survey_year, 
             institution_size)%>%
    summarise_all(list(sum), na.rm = TRUE)

dta$total_vawa <- rowSums(dta[ , colnames(dta[ , 7:ncol(dta)])])

dta_vawa <- dta
```    

# Export

```{r merge}
dim(dta_arrests)
dim(dta_criminal)
dim(dta_vawa)

dta <- left_join(dta_criminal, dta_vawa, by=colnames(dta[ , 1:6]))
dta <- left_join(dta, dta_arrests, by=colnames(dta[ , 1:6]))
dim(dta)

colnames(dta)
```

```{r genvar_more}
dta<-dta%>%
    mutate(total_css = total_vawa + total_criminal + total_arrests)

#create duplicate of numbers
dtarate<-dta[ , 7:ncol(dta)]
colnames(dtarate) <- paste("rate", colnames(dtarate), sep = "_")

#cbind back 
dta<-cbind(dta, dtarate)

#convert to rate 
dta<-dta%>%
    mutate_at(vars(starts_with("rate_")), 
              funs( round(1000* . / institution_size, 2) ) )

```    
              
```{r export}
write_xlsx(dta, 
           'OPE_CSS/ope_css.xlsx', 
           col_names = TRUE)
```

# Prep for merge

```{r export_institution_campus_name}
dta_opeid<-dta%>%
    filter(survey_year==2021)%>%
    select(institution_name, campus_name, opeid, campus_id)%>%
    arrange(institution_name, campus_name)

dim(dta_opeid)

write_xlsx(dta_opeid, 
           'OPE_CSS/opeid.xlsx', 
           col_names = TRUE)
```

```{r filter_dta_to_merge, results='hide'}
# colleges in the analysis 
dtatomerge <- read_excel("CommonDataSet.xlsx", 
                  range = "A3:E1000",
                  col_names = TRUE)%>%
    select(college, year, opeid, campus_id)%>%
    rename(survey_year = year)%>%
    filter(is.na(college)==FALSE)%>%
    
    group_by(college)%>%
    fill(opeid, .direction = "down")%>%
    fill(campus_id, .direction = "down")%>%
    ungroup()

```

```{r merge_prep}
# colnames(dtatomerge)
# colnames(dta)
# dim(dtatomerge)
# dim(dta)

dta <- left_join(dtatomerge, dta, 
                 by=c("opeid", "campus_id", "survey_year"))%>%
    rename(year = survey_year)%>%
    select(-college)
```

```{r export_ope_css_merge_with_cds}
write_xlsx(dta, 
           'OPE_CSS/ope_css_merge_with_cds.xlsx', 
           col_names = TRUE)

```

# Plot tests
```{r JHU_reference}
temp <- read_excel('OPE_CSS/ope_css_merge_with_cds.xlsx', 
                  col_names = TRUE)%>%
    select(-starts_with("..."))%>%
    filter(institution_name=="Johns Hopkins University")%>%
    #filter(campus_id<=2)%>%
    #filter(year==2021)%>%
    select(ends_with("name"), starts_with("rate_total"))

head(temp)
```    

```{r}
##### 0.3 Define color list #####
greycolors <- brewer.pal(7,"Greys")
bluecolors <- brewer.pal(7,"Blues")
greencolors <- brewer.pal(7,"Greens")
orangecolors <- brewer.pal(7,"Oranges")
redcolors <- brewer.pal(7,"Reds")
purplecolors <- brewer.pal(7,"Purples")
divcolors<-brewer.pal(9,"RdYlBu")
qualcolors<-brewer.pal(7,"Set3")

##### 0.4 Define functions for plot #####
hline <- function(y = 0, color = "#CCCCCC") {
    list(
        type = "line",
        x0 = 0,
        x1 = 1,
        xref = "paper",
        y0 = y,
        y1 = y,
        line = list(color = color)
    )
}

vline <- function(x = 0, color = "#CCCCCC") {
    list(
        type = "line",
        x0 = x,
        x1 = x,
        y0 = 0,
        y1 = 1,
        yref = "paper",
        line = list(color = color)
    )
}
```


```{r}
dta<-dta%>%
    mutate(college = institution_name)
```

```{r plot_rate_arrest_rank, results='asis'}

        dtafig<-dta%>%
            
            filter( is.na(institution_size)==FALSE)%>%
            group_by(institution_name)%>%
            mutate(yearlatest=max(year))%>%
            ungroup()%>%
            
            filter(year==yearlatest)
        
        dtafig%>%
            plot_ly(
                y = ~college,
                type="bar", 
                x = ~total_arrests, 
                line = list(color = greencolors[3]), 
                marker = list(color = greencolors[3]),
                hovertemplate = paste0(
                    dtafig$college,       
                    "<br>Number of any arrests (per 1000): ", dtafig$total_arrests))%>%   
            layout(
                title = "", 
                xaxis = list(title = "<br>Number (per 1000)",  
                             #range=c(0, 100), 
                             showgrid = FALSE, showticklabels = TRUE),
                yaxis = list(title = "",
                             tickfont = list(size=9),
                             categoryorder = "total descending"),
                showlegend = FALSE
            )
```

```{r plot_rate_arrest, results='asis'}
        
        dtafig<-dta
        
        dtafig%>%
            plot_ly(
                x = ~year, y = ~total_arrests,
                type="scatter", mode = 'lines+markers', 
                color = ~college,
                hovertemplate = paste0(
                    dtafig$college,       
                    "<br>Year: ", dtafig$year, 
                    "<br>Number of any arrests (per 1000): ", dtafig$total_arrests))%>%
            layout(
                title = "", showlegend = FALSE, 
                shapes = list(vline(2021)),
                yaxis = list(title = "<br>Number (per 1000)",  
                             #range=c(0, 100), 
                             showgrid = FALSE, 
                             showticklabels = TRUE),
                xaxis = list(title = "Year", showgrid = FALSE , range=c(2016, 2024))
            )
```

```{r plot_rate_criminal_rank, results='asis'}
        
        dtafig<-dta%>%
            
            filter( is.na(institution_size)==FALSE)%>%
            group_by(institution_name)%>%
            mutate(yearlatest=max(year))%>%
            ungroup()%>%
            
            filter(year==yearlatest)
        
        dtafig%>%
            plot_ly(
                y = ~college,
                type="bar", 
                x = ~total_criminal, 
                line = list(color = greencolors[3]), 
                marker = list(color = greencolors[3]),
                hovertemplate = paste0(
                    dtafig$college,       
                    "<br>Number of any criminal (per 1000): ", dtafig$total_criminal))%>%   
            layout(
                title = "", 
                xaxis = list(title = "<br>Number (per 1000)",  
                             #range=c(0, 100), 
                             showgrid = FALSE, showticklabels = TRUE),
                yaxis = list(title = "",
                             tickfont = list(size=9),
                             categoryorder = "total descending"),
                showlegend = FALSE
            )
```

```{r plot_rate_criminal, results='asis'}
        
        dtafig<-dta
        
        dtafig%>%
            plot_ly(
                x = ~year, y = ~total_criminal,
                type="scatter", mode = 'lines+markers', 
                color = ~college,
                hovertemplate = paste0(
                    dtafig$college,       
                    "<br>Year: ", dtafig$year, 
                    "<br>Number of any criminal (per 1000): ", dtafig$total_criminal))%>%
            layout(
                title = "", showlegend = FALSE, 
                shapes = list(vline(2021)),
                yaxis = list(title = "<br>Number (per 1000)",  
                             #range=c(0, 100), 
                             showgrid = FALSE, 
                             showticklabels = TRUE),
                xaxis = list(title = "Year", showgrid = FALSE , range=c(2016, 2024))
            )
```

```{r plot_rate_vawa_rank, results='asis'}
        
        dtafig<-dta%>%
            
            filter( is.na(institution_size)==FALSE)%>%
            group_by(institution_name)%>%
            mutate(yearlatest=max(year))%>%
            ungroup()%>%
            
            filter(year==yearlatest)
        
        dtafig%>%
            plot_ly(
                y = ~college,
                type="bar", 
                x = ~total_vawa, 
                line = list(color = greencolors[3]), 
                marker = list(color = greencolors[3]),
                hovertemplate = paste0(
                    dtafig$college,       
                    "<br>Number of any vawa (per 1000): ", dtafig$total_vawa))%>%   
            layout(
                title = "", 
                xaxis = list(title = "<br>Number (per 1000)",  
                             #range=c(0, 100), 
                             showgrid = FALSE, showticklabels = TRUE),
                yaxis = list(title = "",
                             tickfont = list(size=9),
                             categoryorder = "total descending"),
                showlegend = FALSE
            )
```

```{r plot_rate_vawa, results='asis'}
        
        dtafig<-dta
        
        dtafig%>%
            plot_ly(
                x = ~year, y = ~total_vawa,
                type="scatter", mode = 'lines+markers', 
                color = ~college,
                hovertemplate = paste0(
                    dtafig$college,       
                    "<br>Year: ", dtafig$year, 
                    "<br>Number of any vawa (per 1000): ", dtafig$total_vawa))%>%
            layout(
                title = "", showlegend = FALSE, 
                shapes = list(vline(2021)),
                yaxis = list(title = "<br>Number (per 1000)",  
                             #range=c(0, 100), 
                             showgrid = FALSE, 
                             showticklabels = TRUE),
                xaxis = list(title = "Year", showgrid = FALSE , range=c(2016, 2024))
            )
